---
title: "Visualization of Big Index Data And Miscellaneous Functions"
output: rmarkdown::html_vignette
description: >
  This is a manual for the first time installers of maccaindex. You'll understand the 
  concept of Purschasing Power Parity, Big Mac Index, as well as interesting visualization
  of this data to ignite your curiousity. There is a numerical function which might come
  in clutch for your data summarization needs!
vignette: >
  %\VignetteIndexEntry{Visualization of Big Index Data And Miscellaneous Functions}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
editor_options: 
  markdown: 
    wrap: 72
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
library(maccaindex)
library(kableExtra)
library(tibble)
set.seed(123456)
```

## 👋 Introduction

If you have ever been to a Macca before, you probably know what Big Mac
is right...? After all it is a staple menu item that is available in
more than 100 countries around the world (Amica 2020). However, most
economic theories are usually foreign concepts to the average person,
the **Big Mac Index** is of no exception as it was devised on the
premise of **Purchasing Power Parity**, which aims to simplifies the
idea of exchange rates.

## 💭 How does it work?

The Big Mac Index evaluates whether exchange rates are overvalued /
undervalued by comparing the costs of Big Mac in different nations.
Identical commodities ought to be priced the same everywhere; therefore,
variances due to exchange rate fluctuations should be the only factor in
price differences between the burgers. The indicator signals whether a
currency is fairly valued by examining these variations in the Big Mac's
costs against another country as a baseline (For example, the US). If
the cost of a Big Mac differs dramatically from country to country, the
exchange rate may not be where it should be. The index acts as a
distinctive and simple indication to comprehend the valuation of world
currencies and potential market imbalances.

## 🎯 Goal

The `maccaindex` package provides functions for visualizations and
analysis of The Economist's Big Mac Index. We aim to spark the fire of learning economic theories through visual mediums. The package
includes the original dataset for the user to play with, a customization
plotting function, a numerical function to deal with multiple types of
vectors, and a Shiny app that allow users to explore and understand the
Big Mac Index data.

## 🤖 Functionalities

### `macca_raw` dataset

#### Description

This is the raw data of the Big Mac Index, downloaded the data from
`tidytuesdayR`, available for download on their Github repository
[here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-12-22/big-mac.csv).
It contains the following variables:

```{r ,echo=FALSE, message=FALSE, warning=FALSE}
desc <-
  tibble(
    variable = c(
      "date",
      "iso_a3",
      "currency_code",
      "name",
      "local_price",
      "dollar_ex",
      "dollar_price",
      "USD_raw",
      "EUR_raw",
      "GBP_raw",
      "JPY_raw",
      "CNY_raw"
    ),
    class = c(
      "date",
      "character",
      "character",
      "double",
      "double",
      "double",
      "double",
      "double",
      "double",
      "double",
      "double",
      "double"
    ),
    description = c(
      "Date of observation",
      "Three-character [ISO 3166-1 country code][iso 3166-1]",
      "Three-character [ISO 4217 currency code][iso 4217]",
      "Country name",
      "Price of a Big Mac in the local currency",
      "Local currency units per dollar",
      "Price of a Big Mac in dollars",
      "Raw index, relative to the US dollar",
      "Raw index, relative to the Euro",
      "Raw index, relative to the British pound",
      "Raw index, relative to the Japanese yen",
      "Raw index, relative to the Chinese yuan"
    )
  )
kbl(desc, table.attr = 'data-quarto-disable-processing="true"') |>
  kable_styling(full_width = FALSE) |>
  column_spec(1, width = "2cm", border_right = T) |>
  column_spec(2, width = "3cm", border_right = T)
```

**Using the data**: 

To access the data in as an object in the local
environment, load it using the `data()` function. The object is named
`macca_raw`.

**Examples**: 

```{r message=FALSE, warning=FALSE}
data("macca_raw", package = "maccaindex")
macca_raw
```

### `magicwand_df(vectorx, vectory)` function

#### Description

This function does a multitude of calculations and results, depending on
the amount of inputs provided and the types of those inputs. 

**These are the capabilities:**

-   If `vectorx` is a numerical vector of at least 5 values, the function
    returns a data frame with two columns:

    -   Names: the names of the values (Minimum, Lower Hinge, Q1,
        Median, Q3, Upper Hinge, Maximum)

    -   Values: the corresponding summary values

-   If `vectorx` is a character or factor vector, the function returns a
    data frame with two columns:

    -   Values: the unique values in vectorx

    -   Frequencies: the frequencies of each value

-   If both `vectorx` and `vectory` are numerical vectors of the same size,
    the function returns a data frame with one column:

    -   Coefficient: the correlation coefficient between vectorx and
        vectory

-   If one of `vectorx` or `vectory` is numerical and the other is a
    character or factor vector, the function returns a data frame with
    two columns:

    -   Levels: the unique levels in the character/factor vector

    -   Mean: the mean value of the numerical vector for each level

-   If both `vectorx` and `vectory` are character/factor vectors, the
    function returns a data frame representing the cross-tabulation of
    `vectorx` and `vectory`
    
-   Invalid inputs for `vectorx` and `vectory` will be reported with an error                accordingly.

**Examples**: 

```{r}
# Single argument functionality for numerical vector
vectorx <- c(1, 2, 3, 4, 5)
result <- magicwand_df(vectorx)
print(result)
```

```{r}
# Single argument functionality for character/factor vector
vectorx <- c("apple", "banana", "apple", "orange", "banana", "orange")
result <- magicwand_df(vectorx)
print(result)
```

```{r}
# Two-argument functionality for numerical vectors
vectorx <- c(1, 2, 3)
vectory <- c(4, 5, 6)
result <- magicwand_df(vectorx, vectory)
print(result)
```

```{r}
# Two-argument functionality for numerical and character/factor vectors
vectorx <- c(1, 2, 1, 2, 1)
vectory <- c("apple", "banana", "apple", "banana", "banana")
result <- magicwand_df(vectorx, vectory)
print(result)
```

```{r}
# Two-argument functionality for character/factor vectors
vectorx <- c("apple", "banana", "apple", "orange", "banana")
vectory <- c("red", "yellow", "red", "yellow", "red")
result <- magicwand_df(vectorx, vectory)
print(result)
```

### `maccapp()` function

#### Description

This function launches the Shiny app that visualizes the different aspects of the Big Mac Index data. This function accepts zero parameter.

To run the app, just simply write the function as it is.

**Example**: 

```{r , eval=FALSE}
# Run this code to launch the app
maccapp()
#> Listening on http://127.0.0.1:xxxx .... the app window should pop up
```

### `macca_inflation_plot("inflachoice")` function

#### Description

This function generates an interactive plot of inflation rates of all countries from 2011-2020 based on the Big Mac Index. The parameter provided must be a string, which then result in the appropriate currency usage depending on the input. 

To plot for US prices, you can input `usd`, `us dollar`, or `us` (case-insensitive).

**Example**:

```{r , eval=FALSE}
# Plotting inflation rate in terms of US prices
macca_inflation_plot("usd")
#> An interactive plot of the inflation rates over time for different countries, US price.

# Plotting inflation rate in terms of local prices
macca_inflation_plot("localprices")
#> An interactive plot of the inflation rates over time for different countries, local price.
```

